---
title: "Minimal Example--Issues Passing dpylr Variable Names"
author: "Randi Garcia"
date: "6/23/2016"
output: html_document
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, eval=FALSE)
```

## Original Problem

Below is a broken function for mapping services needed. 

```{r}
library(mosaic)
load("~/cvc-workshop/perpnew.Rda")

mapPriority <- function(data, sneed = "Medical", vari, colr = "red"){
  
  propneed <- data %>%
    filter(state != "N/") %>%
    group_by(state) %>%
    summarise(total = n(), need = sum(vari == "(1) Yes")) %>%
    mutate(prop = need/total)

  mUSMap(propneed, key="state", fill = "prop")+
    scale_fill_continuous(low="white", high = colr)+
    ggtitle(paste("Proportion of State", sneed, "Need"))

}

mapPriority(data = perpnew, sneed = "Medical", vari = "FU6P1", colr = "red")
```

Notice that even if we change the variable (`FU6P1` to `FU6P2`), the map doesn't change. 

```{r}
mapPriority(data = perpnew, sneed = "Medical", vari = "FU6P2", colr = "red")
```

But if we add in the `mutate_` function, it does change as expected.

```{r}
mapPriority <- function(data, sneed = "Medical", vari, colr = "red"){
  
  propneed <- data %>%
    mutate_(vari=vari) %>%
    filter(state != "N/") %>%
    group_by(state) %>%
    summarise(total = n(), need = sum(vari == "(1) Yes")) %>%
    mutate(prop = need/total)

  mUSMap(propneed, key="state", fill = "prop")+
    scale_fill_continuous(low="white", high = colr)+
    ggtitle(paste("Proportion of State", sneed, "Need"))

}


mapPriority(data = perpnew, sneed = "Medical", vari = "FU6P5", colr = "red")
mapPriority(data = perpnew, sneed = "Medical", vari = "FU6P2", colr = "red")
```

The title change and map color change now work fine. 

The issue, as I understand it, is with getting summarise() to recogize "vari" as a variable name in "data". The function summarise() is taking a standard function, n(), and a nonstandard function, sum(). Hypothetically, if we could use summarise_(sum(vari)) and summarise(n()) separately, we'd be OK, but we cannot do that because we want to perform those two function on the same tbl. 


## Minimal Example with KidsFeet Data
Because the data from the original example isn't publically available, so we'll make a minimal reproducible example. 

```{r, eval=TRUE}
library(mosaic)

domRights <- function(df, var){
  
  df %>%
    group_by(sex) %>%
    summarise(sampsize = n(), rights = sum(var == "R"))
}
```

Here are some simple tallies to see what we would expect for number of right bigger feet and right dominate hand for boys and girls separately. 

```{r, eval=TRUE}
#for expected number of rights
tally(~biggerfoot|sex, data=KidsFeet)
tally(~domhand|sex, data=KidsFeet)

```

But here is what we get using the broken function above. 

```{r, eval=TRUE}
domRights(KidsFeet, "biggerfoot")
domRights(KidsFeet, "domhand")
```

But it's fixed if we add the `mutate_` function.

```{r, eval=TRUE}
domRights <- function(df, var){
  
  df %>%
    mutate_(var=var) %>%
    group_by(sex) %>%
    summarise(sampsize = n(), rights = sum(var == "R"))
  
}

domRights(KidsFeet, "biggerfoot")
domRights(KidsFeet, "domhand")
```


